---
output: 
  html_document: 
    keep_md: yes
---
# Assignment 1
Load Libraries

```{r}
library(knitr)
opts_chunk$set(echo = TRUE)
library(plyr)
library(lubridate)
library(ggplot2)
```

## Loading and preprocessing the data


```{r}
url <- "https://github.com/tadhg24/RepData_PeerAssessment1/raw/master/activity.zip"

# If zip file does not exist then download
if(!file.exists("repdata-data-activity.zip"))
{
        # Download data from Coursera Site to Working Directory
        download.file(url, destfile = "repdata-data-activity.zip")      
}

# If txt file does not exist then unzip from downloaded file
if(!file.exists("activity.csv"))
{
        # Unzip Data 'exdata-data-household_power_consumption.zip' into working directory
        unzip("repdata-data-activity.zip")      
}

Activity <- read.csv("activity.csv", header = TRUE)

# Convert Date POSIXct
Activity$date <- ymd(Activity$date)

# Activity Information
summary(Activity)
```


## What is mean total number of steps taken per day?

1 Summarise steps by day
```{r}
Activity_Dy <- ddply(Activity, ~date, summarise,
                 Steps_Sum = sum(steps, na.rm = T))
Activity_Dy
```

2 Histogram of Steps/Day
```{r}
g <- ggplot(Activity_Dy, aes(Steps_Sum))
g + geom_histogram()
```

3 Mean & Median of the Steps_Sum / Day
```{r}
mean(Activity_Dy$Steps_Sum)
median(Activity_Dy$Steps_Sum)
```


## What is the average daily activity pattern?

Summarise average steps by interval
```{r}
Activity_Int <- ddply(Activity, ~interval, summarise,
                     Steps_Avg = mean(steps, na.rm = T))
```

1 Average steps by interval
```{r}
g <- ggplot(Activity_Int, aes(interval, Steps_Avg))
g + geom_line()
```

2 5-minute interval with max steps
```{r}
subset(Activity_Int$interval, Activity_Int$Steps_Avg == max(Activity_Int$Steps_Avg, na.rm = T))
```


## Imputing missing values


1 Total number of missing values
```{r}
MissingValuesSteps <- sum(is.na(Activity$steps))
MissingValuesSteps
MissingValuesDate <- sum(is.na(Activity$date))
MissingValuesDate
MissingValuesInterval <- sum(is.na(Activity$interval))
MissingValuesInterval
```

2/3 Replace NA with average steps for that interval
```{r}
Activity_New <- Activity

for(i in 1:nrow(Activity_New))
{
        int <- Activity_New$interval[i]
        
        if(is.na(Activity_New$steps[i]) == TRUE)
        {
                Activity_New$steps[i] <- Activity_Int$Steps_Avg[Activity_Int$interval == int]
        }
}
```

4 Histogram of total steps / day
```{r}
g <- ggplot(Activity_New, aes(steps))
g + geom_histogram()
```
```{r}
Activity_New_Dy <- ddply(Activity_New, ~date, summarise,
                     Steps_Sum = sum(steps, na.rm = T))

# Mean
mean(Activity_New_Dy$Steps_Sum)
# Median
median(Activity_New_Dy$Steps_Sum)
```

The Mean & Median are now equal and have increased from before


## Are there differences in activity patterns between weekdays and weekends?

1 Create factor for weekdays & weekend
```{r}
wkdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

Activity_New$Wkday <- factor((weekdays(Activity_New$date) %in% wkdays),
                             levels = c(TRUE, FALSE),
                             labels = c("weekday", "weekend"))

Activity_New_Int <- ddply(Activity_New, .(interval,Wkday), summarise,
                      Steps_Avg = mean(steps, na.rm = T))
```

2 Panel plot results
```{r}
g <- ggplot(Activity_New_Int, aes(interval, Steps_Avg))
g + geom_line() + facet_wrap(~ Wkday, nrow = 2, ncol = 1)
```